% Official and up-to-date UC Berkeley guidelines on
% dissertation formatting can be found here:
% https://grad.berkeley.edu/academic-progress/dissertation/

%%%%%%%%%%%%%%%%%%%%
%%%%% PACKAGES %%%%%
%%%%%%%%%%%%%%%%%%%%

\documentclass{ucbthesis}
%\usepackage{biblatex}
\usepackage{natbib} % use natbib over biblatex to be able to use GSAB .bst
\usepackage{textcomp,marvosym}
\usepackage{amsmath,amssymb}
\usepackage{rotating}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage[hidelinks]{hyperref}
\urlstyle{same}
\usepackage{threeparttable}
\usepackage[font=footnotesize,format=plain,labelfont=bf,up,textfont=up]{caption} % change figure caption style
\usepackage{color,colortbl}
\usepackage{bibentry}
\bibliographystyle{gsabull}

% To compile this file, run "latex thesis", then "biber thesis"
% (or "bibtex thesis", if the output from latex asks for that instead),
% and then "latex thesis" (without the quotes in each case).

% Double spacing, if you want it. Do not use for the final copy.
% \def\dsp{\def\baselinestretch{2.0}\large\normalsize}
% \dsp

% If the Grad. Division insists that the first paragraph of a section
% be indented (like the others), then include this line:
% \usepackage{indentfirst}

\addtolength{\abovecaptionskip}{\baselineskip}

% Not sure why, but this is required by the style class
\newtheorem{theorem}{Jibberish}

% set hyphenation rules if words are not hyphenated properly
%\hyphenation{mar-gin-al-ia}

%%%%%%%%%%%%%%%%%%%
%%%%% SYMBOLS %%%%%
%%%%%%%%%%%%%%%%%%%

\newcommand{\degreesC}{$^{\circ}$C\xspace}
\newcommand{\degrees}{$^{\circ}$\xspace}
\newcommand{\dC}{$\delta^{13}$C\xspace}
\newcommand{\dO}{$\delta^{18}$O\xspace}
\newcommand{\SrSr}{$^{87}$Sr/$^{86}$Sr\xspace}
\newcommand{\permil}{\textperthousand\xspace}
\newcommand{\dsil}{$d$\xspace}
\newcommand{\UPb}{$^{206}$Pb/$^{238}$U\xspace}
\newcommand{\pCOtwo}{$p$CO$_{2}$\xspace}
\newcommand{\COtwo}{CO$_{2}$\xspace}
\newcommand{\SI}{\textit{Supporting Information}\xspace}
\newcommand{\MM}{\textit{Materials and Methods}\xspace}

\definecolor{Yellow}{rgb}{1,1,0.35}

%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% FRONT MATTER %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\nobibliography* % enigmatically required to use bibentry

\title{Planetary cooling, tectonics, and weathering from 1 billion years ago to the present}

\author{Yuem Park}
\degreesemester{Fall}
\degreeyear{2020}
\degree{Doctor of Philosophy}
\chair{Professor Nicholas L. Swanson-Hysell}
\othermembers{Professor Daniel A. Stolper \\
			  Professor Seth Finnegan}
% For a co-chair who is subordinate to the \chair listed above
% \cochair{Professor Benedict Francis Pope}
% For two co-chairs of equal standing (do not use \chair with this one)
% \cochairs{Professor Richard Francis Sony}{Professor Benedict Francis Pope}
\numberofmembers{3}
\field{Earth \& Planetary Science}
% This is optional (default is Berkeley)
% \campus{Berkeley}

\maketitle
% Delete (or comment out) the \approvalpage line for the final version.
\approvalpage
\copyrightpage

\include{abstract}

\begin{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% TOC AND LISTS %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%

\tableofcontents
\clearpage
\listoffigures
\clearpage
\listoftables

%%%%%%%%%%%%%%%%%%%
%%%%% PREFACE %%%%%
%%%%%%%%%%%%%%%%%%%

\begin{preface}
\phantomsection % make the ToC links work
\addcontentsline{toc}{chapter}{Preface} % add to ToC

The material within the chapters of this dissertation are largely taken from previously published (or soon to be published) articles. The following text identifies these articles, and provides some additional context about how they relate to one another.

\subsubsection*{Chapter 1 - Evaluating the relationship between the area and latitude of large igneous provinces and Earth's long-term climate state}

\noindent
\bibentry{Macdonald2019a}

\medskip

\noindent
\bibentry{Park2019a}

\bigskip

In \citet{Macdonald2019a}, we developed a workflow built on pyGPlates (software for the analysis of paleogeographic models) that allowed us to calculate the area/length of paleogeographically reconstructed polygons/lines within different latitude bands through time. By pairing this workflow with a database of ophiolite-bearing sutures and a paleogeographic model, we were able to identify a strong correlation between arc-continent collisions in the tropics (measured as active suture length in the tropics) and times of glacial climate over the past $\sim$520~m.y. This correlation led to our proposal that Earth's climate state is set primarily by global weatherability, which changes with the latitudinal distribution of arc-continent collisions. Similarly, it has been proposed elsewhere that the eruption/drift of large igneous provinces into the tropics has driven global cooling. The pyGPlates workflow that we developed provided a means through which we could evaluate this hypothesis. \citet{Park2019a} builds upon the zonal large igneous province area analysis presented in the supplementary materials for \citet{Macdonald2019a} by more rigorously developing parameterizations of large igneous province erosion and exploring several geologically reasonable large igneous province post-emplacement scenarios.

\citet{Park2019a} is currently available as a preprint on EarthArXiv, but a slightly modified version of the article is scheduled to be formally published within 2020 as Chapter 7 of AGU Geophysical Monograph Series 255, titled \textit{Large Igneous Provinces: A Driver of Global Environmental and Biotic Changes}, edited by Richard E. Ernst, Alexander J. Dickson, and Andrey Bekker.

Code used to perform the analyses presented in these articles are available on GitHub at \url{https://github.com/Swanson-Hysell-Group/Arc_Continent_Analysis} and \url{https://github.com/Swanson-Hysell-Group/2020_large_igneous_provinces}, or on Zenodo at \url{https://doi.org/10.5281/zenodo.2636731} and \url{https://doi.org/10.5281/zenodo.3981262}.

\subsubsection*{Chapter 2 - Emergence of the Southeast Asian islands as a driver for Neogene cooling}

\noindent
\bibentry{Park2020b}

\bigskip

This article was the result of a collaboration between Godd\'eris' and Swanson-Hysell's research groups supported by the France-Berkeley Fund. Broadly speaking, the collaboration was intended to investigate whether numerical Earth system models quantitatively support the hypothesis put forward in \citet{Macdonald2019a} and \citet{Swanson-Hysell2017a} that tropical arc-continent collisions set Earth's climate state. While the project that was initially proposed sought to investigate whether arc-continent collisions during the Taconic Orogeny led to glaciation in the Ordovician \citep{Swanson-Hysell2017a}, the project changed trajectory to instead investigate more recent changes in the Southeast Asian islands, when constraints on paleogeography are more robust. The collaboration between the two research groups is ongoing, with Maffre, a former Ph.D. candidate in Godd\'eris' group, currently working as a post-doctoral scholar in Swanson-Hysell's group at the time of writing.

The code for the GEOCLIM model used in this study can be found at \url{https://github.com/piermafrost/GEOCLIM-dynsoil-steady-state/releases/tag/v1.0}. The code that generated the inputs and analyzed the output of the GEOCLIM model can be found at \url{https://github.com/Swanson-Hysell-Group/2020_Southeast_Asian_Islands} or \url{https://doi.org/10.5281/zenodo.4021653}.

\subsubsection*{Chapter 3 - The lead-up to the Sturtian Snowball Earth: Neoproterozoic chemostratigraphy time-calibrated by the Tambien Group of Ethiopia}

\noindent
\bibentry{MacLennan2018a}

\medskip

\noindent
\bibentry{Park2020a}

\bigskip

Chronologically, the work that lead to \citet{MacLennan2018a} and \citet{Park2020a} largely preceded the projects presented in Chapters 1 and 2. In \citet{MacLennan2018a}, we presented the initial geochronologic results that placed temporal constraints on large excursions in the marine \dC record and the initiation of the Sturtian glaciation in a short-format journal. \citet{Park2020a} built upon this work by: thoroughly documenting the geology of the Tambien Group in northern Ethiopia, presenting additional geochronologic results, assessing diagenesis in carbonate samples, presenting the complete \dC and \SrSr data from the Tambien Group, developing a composite Tonian-Cryogenian \dC and \SrSr chemostratigraphic record, and proposing a model for the initiation of the Sturtian glaciation based on the composite \SrSr chemostratigraphic record, a simple global weathering model, and analyses of the timing and paleolatitude of large igneous province eruptions and arc accretion events. As in \citet{Swanson-Hysell2017a} and \citet{Macdonald2019a}, the model for the initiation of the Sturtian glaciation proposed in \citet{Park2020a} called upon arc-continent collisions in the tropics, and was part of the motivation to initiate a collaboration with Godd\'eris' research group.

Code used to perform the analyses presented in these articles is available on GitHub at \url{https://github.com/Swanson-Hysell-Group/2019_Tambien_Group}, or on Zenodo at \url{https://doi.org/10.5281/zenodo.3403180}.

\subsubsection*{Chapter 4 - Tonian paleomagnetism from South China permits either an inclusive Rodinia or Bitter Springs true polar wander, but not both}

\noindent
\textit{PRE-PRINT CITATION WILL GO HERE}
%\bibentry{Park2020c}

\bigskip

Although \citet{Park2020a} called upon arc-continent collisions in the tropics as a critical tectonic boundary condition that allowed for the Sturtian glaciation to take place, a lack of paleomagnetic data to constrain the paleolatitude and configuration of tectonic blocks during the lead up to the glaciation hampered the robust defense of this hypothesis. \citet{Park2020c} develops paleomagnetic data to better constrain the location of one of these tectonic blocks, South China, in the Tonian. South China is a particularly important block to investigate with paleomagnetic data in the context of the Sturtian glaciation because arc-continent collisions are hypothesized to have occurred both within and along the margin of the block during the Tonian. Furthermore, sediments from which paleomagnetic data can be developed exist in South China that span a hypothesized major global tectonic event (Bitter Springs Stage true polar wander) that, if real, has far-reaching implications for the paleogeography of the time.

\citet{Park2020c} is currently available as a preprint on EarthArXiv, and has been submitted to the \textit{Journal of Geophysical Research: Solid Earth} for review.

Code used to perform the analyses presented in these articles is available on GitHub at \url{xxx}, or on Zenodo at \url{xxx}.

\end{preface}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% ACKNOWLEDGEMENTS %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{acknowledgements}
\phantomsection % make the ToC links work
\addcontentsline{toc}{chapter}{Acknowledgements} % add to ToC
Bovinely invasive brag; cerulean forebearance.
Washable an acre. To canned, silence in foreign.
Be a popularly. A as midnight transcript alike.
To by recollection bleeding. That calf are infant. In clause.
Buckaroo loquaciousness?  Aristotelian!
Masterpiece as devoted. My primal the narcotic. For cine?
In the glitter. For so talented. Which is confines cocoa accomplished.
Or obstructive, or purposeful.
And exposition? Of go. No upstairs do fingering.
\end{acknowledgements}

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%
%%%%% CHAPTERS %%%%%
%%%%%%%%%%%%%%%%%%%%

\pagestyle{headings}

% (Optional) \part{First Part}

\include{chap-LIPs}
\include{chap-SEAIs}
\include{chap-Tambien}

%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% BIBLIOGRAPHY %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{references}

%%%%%%%%%%%%%%%%%%%%
%%%%% APPENDIX %%%%%
%%%%%%%%%%%%%%%%%%%%

\appendix
\include{SI-SEAIs}
\include{SI-Tambien}

\end{document}
